---
title: "Depth Profiles"
author: "Lilah McCormick"
date: '2022-06-30'
output: 
  html_document:
    toc: yes
    toc_float: yes 
    theme: default 
    highlights: monotone
---

# Before Starting: load libraries
```{r = load_packages, messages = FALSE}
library(lubridate)
library(tidyverse)
library(plotly)
library(DT)
library(glue)
```


# Task #1: Read the file into a dataframe.   
### Step 1: load the data.
```{r read_txt_file}
depth_data <- read_tsv("Depth_Asgeir.txt")
header <- head(depth_data)
class(depth_data) #checks that the file has been read as a dataframe.
```
    
# Task #2: Convert the datetime column from character to POSIXct.   
### Step 1: Outline changes that need to happen.    
Format of datetime entry in data:    
`m/dd/yy hh:mm:ss`
`parse_datetime()` expects: 
`yyyy/mm/dd hh:mm:ss`

### Step 2: Figure out how we need to manipulate the given date using the date from the first entry  
```{r practice_datetime}
line_one <- "8/23/18 22:46:01"
mdy_hms(line_one)
```

### Step 3 Pipe on Header
```{r header_datetime_pipe}
header_datetime <- read_tsv("Depth_Asgeir.txt") %>% 
  head() %>% 
  mutate(DateTime = mdy_hms(`Date time`), .keep = "unused")  #fixes datetime 
  
```

### Step 4 Pipe on full dataset 
```{r = fix_datetimes}
depth_data <- read_tsv("Depth_Asgeir.txt") %>% 
  mutate(DateTime = mdy_hms(`Date time`), .keep = "unused")#fixes datetime
```

### Step 4: Apply this set of functions to the whole dataset. 
```{r full_dt_conversion, message = FALSE}
datetime_col <- pull(depth_data, "Date time") #creates a vector with all datetimes
UTC_vec <- mdy_hms(datetime_col) #changes into UTC
UTC_col <- data.frame(UTC_vec) #changes it back into a df 
depth_data_UTC <- mutate(depth_data, UTC_col, .before = "Date time") #Adds UTC datetime column in before old date time column 
depth_data_UTC <- select(depth_data_UTC, -c("Date time")) #delete old datetime column
depth_data_UTC <- rename(depth_data_UTC, DateTime = UTC_vec) #renames the datetime column 
```


### Step 5: Print header to show the new table
```{r = print_new_header}
head(depth_data_UTC)
```

# Task #3 Make Figures
1. Depth profile for the whole deployment
1. Depth profile for 1 day 
1. Depth profile for 1 dive 

### Step 1: Make a plot for the whole deployment   
```{r plot_deployment}
deployment_plot <- ggplot(depth_data_UTC, aes(x = DateTime, y = Depth)) + 
  geom_line() +
  scale_y_reverse() + #makes depth right orientation 
  labs(title = "Asgeir Depth Profile 2018",
       x = "Time",
       y = "Depth (m)") + #changes labels on axes and title 
  theme_bw()  #white background 
```

### Step 2: Make a plot for 1 day  

I'm a little confused on this one - This is a 24 hour period, but it doesn't start at midnight, and I think we might need to know the timezone for this one.
```{r plot_day}
start_day <- ymd_hms('2018-08-24 00:00:00')
end_day <- start_day + days(1)
one_day <- filter(depth_data_UTC, DateTime >= start_day, DateTime <= end_day)
day_plot<- ggplot(one_day, aes(x = DateTime, y = Depth)) + #filters datetime to 24 hour period
  geom_line() +
  scale_y_reverse() + #makes depth right orientation 
  labs(title = "Asgeir 24 Hour Depth Profile 2018",
       x = "Time",
       y = "Depth (m)") + #changes labels on axes and title 
  theme_bw()  #white background 
ggplotly(day_plot) #makes plot interactive 
```

### Step 3: Make a plot for 1 dive
Also have to figure out some time zone stuff on this one. 
```{r plot_one_dive}
start_dive_time <- ymd_hms('2018-08-24 08:26:01') #selected times from interactive day plot 
end_dive_time <- ymd_hms('2018-08-24 08:49:04')
one_dive <- filter(depth_data_UTC, DateTime >= start_dive_time, DateTime <= end_dive_time)
dive_plot <- ggplot(one_dive, aes(x = DateTime, y = Depth)) + #filters datetime to 24 hour period
  geom_line() +
  scale_y_reverse() + #makes depth right orientation 
  labs(title = "Asgeir Single Dive Depth Profile 2018",
       x = "Time",
       y = "Depth (m)") + #changes labels on axes and title 
  theme_bw()  #white background 
#ggplotly(dive_plot)
dive_plot
```



