---
title: "Depth Profiles"
author: "Lilah McCormick"
date: '2022-06-30'
output: 
  html_document:
  toc: yes
  toc_float: yes 
  theme: paper 
  highlights: monotone
---
# Questions for Max:
1. Do I need to install a package at the beginning of every RMarkdown file? eg. lubridate  
1. Do I need to call the library at the beginning of every chunk of code?   

# Task #1: Read the file into a dataframe. 
### Step 1: load the data. **Note:** Make sure you spell the name of the file correctly :(
```{r = read_txt_file}
depth_data <- read.delim("Depth_Asgeir.txt")
head(depth_data)
class(depth_data) #checks that the file has been read as a dataframe.
```

# Task #2: Convert the datetime column from character to POSIXct. 
### Step 1: Outline changes that need to happen.  
Format of datetime entry in data:  
`m/dd/yy hh:mm:ss`
`parse_datetime()` expects: 
`yyyy/mm/dd hh:mm:ss`

### Step 2: Figure out how we need to manipulate the given date using a practice character vector 
```{r = practice_datetime}
library(lubridate)
parse_date_time("8/30/22", "m/%d/%y")
```



